---
title: "Historical U.S.A. Boundaries"
author: "Lincoln Mullen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

The [USAboundaries](https://github.com/ropensci/USAboundaries/) package provides an R wrapper around the boundary data for the United States of America in the [Atlas of Historical County Boundaries](http://publications.newberry.org/ahcbp/) from the Dr. William M. Scholl Center for American History and Culture at the Newberry Library. This package provides lower resolution shapefiles; for higher resolutions, see the original project page.

```{r}
library(USAboundaries)
library(maptools)
library(sp)
library(ggplot2)
```

You can return either states (the default) or counties. This plot shows the states in 1844.

```{r fig.cap="The United States in 1844"}
us_1844 <- us_boundaries(as.Date("1844-07-04"))
plot(us_1844)
```

This plot shows the counties in Virginia in 1844.

```{r fig.cap="Counties in Virginia in 1844"}
va_1844 <- us_boundaries(as.Date("1844-07-04"), states = "Virginia", 
                         type = "county")
plot(va_1844)
```

By default object returned is of class `sp` from the [sp](http://cran.rstudio.org/web/packages/sp/) package. You can also return a data frame suitable for plotting in [ggplot2](http://cran.rstudio.org/web/packages/ggplot2/).

```{r fig.cap="United States in 1790"}
us_1790 <- us_boundaries(as.Date("1790-07-04"), format = "df")
ggplot(data = us_1790) +
  geom_map(map = us_1790,
           aes(x = long, y = lat, group = group, map_id = id),
           fill = "white", color = "black", size = 0.2) +
  theme_minimal()
```

See the documentation for the `us_boundaries()` function for more information.
