---
title: "Historical and Contemporary Boundaries of the United States of America"
author: "Lincoln Mullen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

This R package includes contemporary state, county, congressional district, and zip code tabulation area boundaries for the United States of America, as well as historical boundaries from 1629 to 2000 for states and counties from the Newberry Library's [Atlas of Historical County Boundaries](http://publications.newberry.org/ahcbp/). This package provides a set of functions, one for each of the types of boundaries that are available. These functions have a consistent interface described below.

## Types of data

There are four kinds of boundaries provided by this package, each with its own function:

- State: `us_states()`
- Counties: `us_counties()`
- Congressional districts: `us_congressional()`
- Zip code tabulation areas: `us_zipcodes()`

In addition, the `us_boundaries()` function provides a wrapper around all of those functions for backward compatibility. But in general you are better off using the more specific functions.

## Historical vs contemporary boundaries

The `map_date = ` argument controls whether the functions return historical or contemporary boundary information. Leaving the argument `NULL` will always return contemporary boundaries.

```{r}
library(USAboundaries)
library(sp) # for plotting

states <- us_states()
plot(states)
title("Contemporary state boundaries")
```

Passing a date as a character vector (or any object coercible to a date object with `as.Date()`) will return the boundaries for that day.

```{r}
states_1790 <- us_states("1790-07-04")
plot(states_1790)
title("State boundaries in 1790")
```

Historical data is only available for states and counties. The functions `us_congressional()` and `us_zipcodes()` only offer contemporary boundaries.

## Selection certain states

For any function, pass a character vector of state names to the `states =` argument to return only those states or territories.

```{r}
counties <- us_counties(states = c("South Carolina", "North Carolina"))
plot(counties)
title("Contemporary county boundaries in North and South Carolina")
```

## Resolution

For certain functions, more or less detailed boundary information is available by passing an argument to the `resolution =` argument.

```{r}
counties_va <- us_counties(states = "Virginia")
plot(counties_va)
title("Contemporary county boundaries in Virginia")

counties_va_highres <- us_counties(states = "Virginia", resolution = "500k")
plot(counties_va_highres)
title("Higher resolution contemporary county boundaries in Virginia")
```
